---
title: "flat_first.Rmd for working package"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- Run this 'development' chunk -->
<!-- Store every call to library() that you need to explore your functions -->

```{r development, include=FALSE}
library(testthat)
```

<!--
 You need to run the 'description' chunk in the '0-dev_history.Rmd' file before continuing your code there.

If it is the first time you use {fusen}, after 'description', you can directly run the last chunk of the present file with inflate() inside.
--> 

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```



# charge_stations_aspe

<!--
Create a chunk for the core of the function

- The chunk needs to be named `function` at least
- It contains the code of a documented function
- The chunk can also be named `function-my_median` to make it easily
findable in your Rmd
- Let the `@examples` part empty, and use the next `examples` chunk instead to present reproducible examples

After inflating the template

-  This function code will automatically be added in a new file in the "R/" directory
-->

```{r function}
#' charge_stations_aspe
#'
#' @param dpt Vecteur avec les départements à charger
#' @param code_agence Vecteur avec les codes agences selon https://id.eaufrance.fr/nsa/447
#'
#' @return
#' Liste des stations ciblées
#' @export
#'
#' @examples
charge_stations_aspe <- function(dpt, code_agence = c("04")) {

# URL de base de l'API
base_url <- "https://hubeau.eaufrance.fr/api/v1/etat_piscicole/stations"

dpt0<-paste(dpt, sep="", collapse = ",")
code_bassin0<-paste(code_agence, sep="", collapse = ",")

# Paramètres de base de la requête
params <- list(
  code_departement = dpt0, # Codes INSEE des départements
  code_bassin = code_bassin0,                    # Code du bassin
  format = "json",                       # Format de la réponse
  size = 100,                             # Taille de la page (nombre de résultats par requête)
fields = "code_station,libelle_station,coordonnee_x_station,coordonnee_y_station,code_point_prelevement_aspe,code_point_prelevement,longitude,latitude"
  )

# Initialiser les variables pour la pagination
all_stations <- list() # Liste pour stocker toutes les stations
page <- 1              # Commencer à la première page
total_results <- Inf   # Nombre total de résultats à récupérer, infini pour commencer

# Boucle pour récupérer toutes les pages de résultats
while ((page - 1) * params$size < total_results) {
  # Ajouter le paramètre de pagination à la requête
  params$page <- page
  
  # Faire une requête GET à l'API
  response <- httr::GET(url = base_url, query = params)
  
  # Vérifier le code de statut
  if (httr::status_code(response) %in% c(200, 206)) {
    # Convertir le contenu JSON de la réponse en liste R
    stations_data <- jsonlite::fromJSON(httr::content(response, as = "text", encoding = "UTF-8"))
    
   # Vérifier si des données sont présentes
    if (!is.null(stations_data$data) && length(stations_data$data) > 0) {
      # Convertir les données en un data.frame
      stations_df <- as.data.frame(stations_data$data)
      
      # Ajouter ce data.frame à la liste totale
      all_stations <- c(all_stations, list(stations_df))
    }    
    # Mettre à jour le nombre total de résultats si disponible
    if (!is.null(stations_data$count)) {
      total_results <- stations_data$count
    } else {
      # Si l'info de total est absente, on suppose qu'on a tout après la première réponse 200 sans nouvelle data
      break
    }
    
    # Afficher une information de progression
    message(paste("Page", page, "de", ceiling(total_results / params$size), "récupérée."))
    
    # Passer à la page suivante
    page <- page + 1
  } else {
    # Afficher un message d'erreur si la requête a échoué
    stop(paste("Erreur:", httr::status_code(response)))
  }
}

# Convertir la liste de stations en data.frame pour analyse ou export
stations_df <- do.call(rbind, lapply(all_stations, as.data.frame))
return(stations_df)
}
```

<!--
Create a chunk with an example of use for your function

- The chunk needs to be named `examples` at least
- It contains working examples of your function
- The chunk is better be named `examples-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This example will automatically be added in the '@examples' part of our function above in the "R/" directory
- This example will automatically be added in the vignette created from this Rmd template
-->

```{r examples}

dpt<-c("44", "56", "22", "35", "49", "53")
liste_stations<-charge_stations_aspe(dpt, code_agence = c("04"))


```

<!--
Create a chunk with a test of use for your function

- The chunk needs to be named `tests` at least
- It contains working tests of your function
- The chunk is better be named `tests-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This test code will automatically be added in the "tests/testthat/" directory
-->

```{r tests}
test_that("charge_stations_aspe works properly and show error if needed", {
  dpt<-c("44")
liste_stations<-charge_stations_aspe(dpt, code_agence = c("04"))
  expect_true(nrow(liste_stations)>0)
})


```

# liste_operations_aspe


```{r function}
#' liste_operations_aspe
#'
#' @param codes_stations Vecteur avec les codes stations dont on veut les opérations.
#' @param code_point_prelevement_aspe Vecteur avec les codes stations dont on veut les opérations
#' @param date_debut Date minimale de création de l'opération en base de données. (format date)
#' @param date_fin Date maximale de création de l'opération en base de données (format date)
#' @param dpt Vecteur avec les départements à charger
#' @param code_agence Vecteur avec les codes agences selon https://id.eaufrance.fr/nsa/447
#' 
#' @return
#' Liste des operations
#' @export
#'
#' @examples
liste_operations_aspe <- function(codes_stations = NULL,
                                  code_point_prelevement_aspe = NULL,
                                  date_debut = NULL,
                                  date_fin = NULL,
                                  dpt = NULL,
                                  code_agence = NULL) {
  # URL de base de l'API
  base_url <- "https://hubeau.eaufrance.fr/api/v1/etat_piscicole/operations"
  
  # Vérification et formatage des paramètres
  if (!is.null(codes_stations)) {
    if (!is.character(codes_stations)) {
      stop("codes_stations doit être un vecteur de caractères.")
    }
    if (length(codes_stations)>200) {
      stop("codes_stations doit comporter moins de 200 valeurs")
    }
    codes_stations0 <- paste(codes_stations, sep = "", collapse = ",")
  } else {
    codes_stations0 <- NULL
  }
  
  if (!is.null(code_point_prelevement_aspe)) {
    if (!is.character(code_point_prelevement_aspe)) {
      stop("code_point_prelevement_aspe doit être un vecteur de caractères.")
    }
    if (length(code_point_prelevement_aspe)>200) {
      stop("code_point_prelevement_aspe doit comporter moins de 200 valeurs")
    }
    code_point_prelevement_aspe0 <- paste(code_point_prelevement_aspe, sep = "", collapse = ",")
  } else {
    code_point_prelevement_aspe0 <- NULL
  }
  
  
  if (!is.null(date_debut)) {
    if (!lubridate::is.Date(lubridate::ymd(date_debut))) {
      stop("date_debut doit être au format date (YYYY-MM-DD).")
    }
    date_debut <- as.character(lubridate::ymd(date_debut))
  }
  
  if (!is.null(date_fin)) {
    if (!lubridate::is.Date(lubridate::ymd(date_fin))) {
      stop("date_fin doit être au format date (YYYY-MM-DD).")
    }
    date_fin <- as.character(lubridate::ymd(date_fin))
  }
  
  if (!is.null(dpt)) {
    if (!is.character(dpt)) {
      stop("dpt doit être un vecteur de caractères.")
    }
    dpt0 <- paste(dpt, sep = "", collapse = ",")
  } else {
    dpt0 <- NULL
  }
  
  if (!is.null(code_agence)) {
    if (!is.character(code_agence))
    {
      stop("code_agence doit être un vecteur de caractères.")
      code_agence <- paste(code_agence, sep = "", collapse = ",")
    }
  } else {
    code_agence <- NULL
  }
  
  # Paramètres de base de la requête
  params <- list(
    code_station = codes_stations0,
    code_point_prelevement_aspe = code_point_prelevement_aspe0,
    # Codes stations
    date_creation_operation_min = if (!is.null(date_debut))
      date_debut
    else
      NULL,
    date_creation_operation_max = if (!is.null(date_fin))
      date_fin
    else
      NULL,
    code_departement = dpt0,
    # Codes INSEE des départements
    code_bassin = code_agence,  # Code du bassin
    format = "json",
    # Format de la réponse
    size = 100,
    # Taille de la page (nombre de résultats par requête)
    fields = "code_operation,date_operation,etat_avancement_operation,
libelle_qualification_operation,code_station,code_point_prelevement,
coordonnee_x_point_prelevement,coordonnee_y_point_prelevement,protocole_peche,
moyen_prospection,operateur_libelle,commanditaire_libelle,date_creation_operation,
date_modification_operation"
  )

    params <- Filter(Negate(is.null), params)

  # Initialiser les variables pour la pagination
  all_stations <- list() # Liste pour stocker toutes les stations
  page <- 1              # Commencer à la première page
  total_results <- Inf   # Nombre total de résultats à récupérer, infini pour commencer
  
  # Boucle pour récupérer toutes les pages de résultats
  while ((page - 1) * params$size < total_results) {
    # Ajouter le paramètre de pagination à la requête
    params$page <- page
    
    # Faire une requête GET à l'API
    response <- httr::GET(url = base_url, query = params)
    
    # Vérifier le code de statut
    if (httr::status_code(response) %in% c(200, 206)) {
      # Convertir le contenu JSON de la réponse en liste R
      stations_data <- jsonlite::fromJSON(httr::content(response, as = "text", encoding = "UTF-8"))
      
      # Vérifier si des données sont présentes
      if (!is.null(stations_data$data) &&
          length(stations_data$data) > 0) {
        # Convertir les données en un data.frame
        stations_df <- as.data.frame(stations_data$data)
        
        # Ajouter ce data.frame à la liste totale
        all_stations <- c(all_stations, list(stations_df))
      }
      # Mettre à jour le nombre total de résultats si disponible
      if (!is.null(stations_data$count)) {
        total_results <- stations_data$count
      } else {
        # Si l'info de total est absente, on suppose qu'on a tout après la première réponse 200 sans nouvelle data
        break
      }
      
      # Afficher une information de progression
      message(paste(
        "Page",
        page,
        "de",
        ceiling(total_results / params$size),
        "récupérée."
      ))
      
      # Passer à la page suivante
      page <- page + 1
    } else {
      # Afficher un message d'erreur si la requête a échoué
      stop(paste("Erreur:", httr::status_code(response)))
    }
  }
  
  # Convertir la liste de stations en data.frame pour analyse ou export
  stations_df <- do.call(rbind, lapply(all_stations, as.data.frame))
  return(stations_df)
}

```

<!--
Create a chunk with an example of use for your function

- The chunk needs to be named `examples` at least
- It contains working examples of your function
- The chunk is better be named `examples-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This example will automatically be added in the '@examples' part of our function above in the "R/" directory
- This example will automatically be added in the vignette created from this Rmd template
-->

```{r examples}

dpt<-c("44", "56")
liste_stations<-charge_stations_aspe(dpt, code_agence = c("04"))
liste_stations<-liste_stations[!is.na(liste_stations$code_station),]
liste_op<-liste_operations_aspe(codes_stations=head(liste_stations$code_station,10), 
                      date_debut=as.Date("2010-01-01"), 
                      date_fin=as.Date("2024-01-01")) 

liste_op2<-liste_operations_aspe(code_point_prelevement_aspe = head(liste_stations$code_point_prelevement_aspe,10),                       date_debut=as.Date("2010-01-01"), 
                      date_fin=as.Date("2024-01-01")) 


```

<!--
Create a chunk with a test of use for your function

- The chunk needs to be named `tests` at least
- It contains working tests of your function
- The chunk is better be named `tests-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This test code will automatically be added in the "tests/testthat/" directory
-->

```{r tests}
test_that("liste_operations_aspe works properly and show error if needed", {
 dpt<-c("53")
liste_stations<-charge_stations_aspe(dpt, code_agence = c("04"))
liste_stations<-liste_stations[!is.na(liste_stations$code_station),]
liste_op<-liste_operations_aspe(codes_stations=head(liste_stations$code_station,10), 
                      date_debut=as.Date("2010-01-01"), 
                      date_fin=as.Date("2024-01-01"))
  expect_true(nrow(liste_op)>0)
})


```

# charge_observations_aspe

<!--
Create a chunk for the core of the function

- The chunk needs to be named `function` at least
- It contains the code of a documented function
- The chunk can also be named `function-my_median` to make it easily
findable in your Rmd
- Let the `@examples` part empty, and use the next `examples` chunk instead to present reproducible examples

After inflating the template

-  This function code will automatically be added in a new file in the "R/" directory
-->

```{r function}
#' charge_observations_aspe
#'
#' @param code_operation Vecteur avec les codes opérations à charger (paramètre obligatoire)
#'
#' @return
#' Liste des stations ciblées
#' @export
#'
#' @examples
charge_observations_aspe <- function(code_operation) {
  # URL de base de l'API
  base_url <- "https://hubeau.eaufrance.fr/api/v1/etat_piscicole/observations.csv"
  
  # Vérification et formatage des paramètres
  if (!is.null(code_operation)) {
    if (!is.character(code_operation)) {
      stop("code_operation doit être un vecteur de caractères.")
    }
    if (length(code_operation)>200) {
      stop("code_operation doit comporter moins de 200 valeurs")
    }
    code_operation <- paste(code_operation, sep = "", collapse = ",")
  } else {
    stop("code_operation doit être non NULL.")
  }
  
  # Paramètres de base de la requête
  params <- list(
    code_operation = code_operation,
    size = 100  # Taille de la page (nombre de résultats par requête)
  )
  
  # Initialiser les variables pour la pagination
  all_stations <- list() # Liste pour stocker toutes les stations
  page <- 1              # Commencer à la première page
  more_data <- TRUE      # Flag pour indiquer s'il y a encore des données à récupérer
  
  # Boucle pour récupérer toutes les pages de résultats
  while (more_data) {
    # Ajouter le paramètre de pagination à la requête
    params$page <- page
    
    # Faire une requête GET à l'API
    response <- httr::GET(url = base_url, query = params)
    
    # Vérifier le code de statut
    if (httr::status_code(response) %in% c(200, 206)) {
      # Obtenir le contenu de la réponse
      content_text <- httr::content(response, as = "text", encoding = "UTF-8", type="text")
      
      # Vérifier si le contenu est vide
      if (nchar(content_text) == 0) {
        more_data <- FALSE  # Arrêter la pagination s'il n'y a plus de données
      } else {
        # Lire le contenu CSV de la réponse avec ; comme séparateur
        stations_data <- read.csv2(text = content_text, sep = ";")
        
        # Vérifier si des données sont présentes
        if (!is.null(stations_data) && nrow(stations_data) > 0) {
          # Ajouter ce data.frame à la liste totale
          all_stations <- c(all_stations, list(stations_data))
          
          # Afficher une information de progression
          message(paste("Page", page, "récupérée. Nombre d'enregistrements:", nrow(stations_data)))
          
          # Passer à la page suivante
          page <- page + 1
        } else {
          # Arrêter la pagination si le data.frame est vide
          more_data <- FALSE
        }
      }
    } else {
      # Afficher un message d'erreur si la requête a échoué
      stop(paste("Erreur:", httr::status_code(response)))
    }
  }
  
  # Convertir la liste de stations en data.frame pour analyse ou export
  if (length(all_stations) > 0) {
    stations_df <- do.call(rbind, lapply(all_stations, as.data.frame))
    stations_df <- unique(stations_df)
  } else {
    stations_df <- data.frame()  # Retourne un data.frame vide si aucune donnée n'est récupérée
  }
  
  return(stations_df)
}

```


```{r examples}

obs<-charge_observations_aspe(code_operation=c("89771","89805"))

```

<!--
Create a chunk with a test of use for your function

- The chunk needs to be named `tests` at least
- It contains working tests of your function
- The chunk is better be named `tests-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This test code will automatically be added in the "tests/testthat/" directory
-->

```{r tests}
test_that("charge_observations_aspe works properly and show error if needed", {
  expect_true(nrow(charge_observations_aspe(code_operation=c("89771")))>0)
})


```

# split_observations_aspe


```{r function}
#' split_observations_aspe
#'
#' @param observations Résultat renvoyé par la fonction charge_observations_aspe
#'
#' @return
#' Liste avec d'un côté les conditions de l'opération et de l'autre les listes faunistiques
#' @export
#'
#' @examples
split_observations_aspe <- function(observations) {

  detail_operation<-unique(observations[,                                 c("code_operation","codes_dispositifs_collecte","libelles_dispositifs_collecte","protocole_peche","objectifs_operation","code_prelevement_elementaire","type_prelevement_elementaire","duree_prelevement_elementaire","numero_passage","type_points","nombre_points","nombre_points_sans_poisson","nombre_points_facies_courant","nombre_points_facies_plat","nombre_points_facies_profond","nombre_points_annexes","nombre_points_berge","nombre_points_chenal","nom_ambiance","longueur_ambiance","largeur_ambiance","profondeur_ambiance","code_facies_ambiance","libelle_facies_ambiance","situation_ambiance","localisation_ambiance","classe_vitesse_courant_ambiance","code_ombrage_ambiance","libelle_ombrage_ambiance","code_granulo_ambiance_dominante","libelle_granulo_ambiance_dominante","code_granulo_ambiance_accessoire","libelle_granulo_ambiance_accessoire","vegetation_ambiance","abondance_cache_vegetation_aquatique_ambiance","abondance_cache_embacles_ambiance","abondance_cache_vegetation_bordure_ambiance","abondance_cache_sous_berge_ambiance","abondance_cache_granulo_ambiance")])
                          
      liste_taxons<-unique(observations[,                                 c("code_operation","code_lot","code_type_lot","libelle_type_lot","effectif_lot","code_taxon","code_alternatif_taxon","nom_commun_taxon","nom_latin_taxon","uri_taxon","taille_min_lot","taille_max_lot","type_longueur_lot","poids_lot_mesure","poids_lot_estime","methode_estimation_poids_lot","code_individu","taille_individu","type_longueur_individu","mesure_individu_reelle","poids_individu_mesure","poids_individu_estime","methode_estimation_poids_individu","sexe_individu","age_individu","codes_pathologies_lot","libelles_pathologies_lot","codes_fractions_pathologies_lot","libelles_fractions_pathologies_lot","uris_fractions_pathologies_lot","importances_pathologies_lot","codes_pathologies_individu","libelles_pathologies_individu","codes_fractions_pathologies_individu","libelles_fractions_pathologies_individu","uris_fractions_pathologies_individu","importances_pathologies_individu")])
  
  return(list(detail_operation=detail_operation,
         liste_taxons=liste_taxons))
}

```


```{r examples}

obs<-charge_observations_aspe(code_operation=c("89771","89805"))
detail<-split_observations_aspe(obs)


```

<!--
Create a chunk with a test of use for your function

- The chunk needs to be named `tests` at least
- It contains working tests of your function
- The chunk is better be named `tests-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This test code will automatically be added in the "tests/testthat/" directory
-->

```{r tests}
test_that("split_observations_aspe works properly and show error if needed", {
  obs<-charge_observations_aspe(code_operation=c("89771","89805"))
detail<-split_observations_aspe(obs)
  expect_true(nrow(detail$detail_operation)>=2)
expect_true(nrow(detail$liste_taxons)>=100)
})


```

<!--
# There can be development actions

Create a chunk with 'development' actions

- The chunk needs to be named `development` or `dev`
- It contains functions that are used for package development only
- Note that you may want to store most of these functions in the 0-dev_history.Rmd file

These are only included in the present flat template file, their content will not be part of the package anywhere else.
-->

```{r development-inflate, eval=FALSE}
# Keep eval=FALSE to avoid infinite loop in case you hit the knit button
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_first.Rmd", vignette_name = NA)
```


# Inflate your package

You're one inflate from paper to box.
Build your package from this very Rmd using `fusen::inflate()`

- Verify your `"DESCRIPTION"` file has been updated
- Verify your function is in `"R/"` directory
- Verify your test is in `"tests/testthat/"` directory
- Verify this Rmd appears in `"vignettes/"` directory
